Graphical Tests for Power Comparison of Competing Designs

### 2012-03-28 experiment-two.tex
p values in tables

\begin{table}[ht]
\begin{center}
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
  \hline
(Intercept) & -2.00 & 0.38 & -5.29 & 0.00 \\ 
  test\_paramdensity & -0.17 & 0.53 & -0.32 & 0.75 \\ 
  test\_paramdotplot & -0.72 & 0.52 & -1.39 & 0.17 \\ 
  test\_paramhistogram & -0.30 & 0.53 & -0.58 & 0.56 \\ 
  d & 1.77 & 0.34 & 5.19 & 0.00 \\ 
  n1 & 0.02 & 0.00 & 9.80 & 0.00 \\ 
  I(n1/n2) & -0.31 & 0.11 & -2.93 & 0.00 \\ 
  test\_paramdensity:d & -0.67 & 0.47 & -1.42 & 0.15 \\ 
  test\_paramdotplot:d & 0.05 & 0.46 & 0.11 & 0.91 \\ 
  test\_paramhistogram:d & -0.27 & 0.48 & -0.56 & 0.57 \\ 
  test\_paramdensity:n1 & -0.00 & 0.00 & -0.40 & 0.69 \\ 
  test\_paramdotplot:n1 & -0.01 & 0.00 & -3.06 & 0.00 \\ 
  test\_paramhistogram:n1 & -0.01 & 0.00 & -2.67 & 0.01 \\ 
  test\_paramdensity:I(n1/n2) & 0.06 & 0.15 & 0.42 & 0.68 \\ 
  test\_paramdotplot:I(n1/n2) & 0.27 & 0.15 & 1.81 & 0.07 \\ 
  test\_paramhistogram:I(n1/n2) & 0.10 & 0.15 & 0.65 & 0.52 \\ 
   \hline
\end{tabular}
\end{center}
\end{table}

>>>>

\begin{table}[ht]
\begin{center}
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
  \hline
(Intercept) & -2.00 & 0.38 & -5.29 & 0.00 \\ 
  test\_paramdensity & -0.29 & 0.53 & -0.55 & 0.58 \\ 
  test\_paramdotplot & -0.72 & 0.52 & -1.39 & 0.17 \\ 
  test\_paramhistogram & -0.29 & 0.53 & -0.54 & 0.59 \\ 
  d & 1.78 & 0.34 & 5.22 & 0.00 \\ 
  n1 & 0.02 & 0.00 & 9.77 & 0.00 \\ 
  I(n1/n2) & -0.31 & 0.11 & -2.94 & 0.00 \\ 
  test\_paramdensity:d & -0.61 & 0.47 & -1.29 & 0.20 \\ 
  test\_paramdotplot:d & 0.04 & 0.46 & 0.08 & 0.94 \\ 
  test\_paramhistogram:d & -0.29 & 0.48 & -0.60 & 0.55 \\ 
  test\_paramdensity:n1 & -0.00 & 0.00 & -0.40 & 0.69 \\ 
  test\_paramdotplot:n1 & -0.01 & 0.00 & -3.04 & 0.00 \\ 
  test\_paramhistogram:n1 & -0.01 & 0.00 & -2.67 & 0.01 \\ 
  test\_paramdensity:I(n1/n2) & 0.09 & 0.15 & 0.59 & 0.55 \\ 
  test\_paramdotplot:I(n1/n2) & 0.27 & 0.15 & 1.81 & 0.07 \\ 
  test\_paramhistogram:I(n1/n2) & 0.09 & 0.15 & 0.62 & 0.53 \\ 
   \hline
\end{tabular}
\end{center}
\caption{correctness}
\end{table}

\begin{table}[ht]
\begin{center}
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std..Error & t.value & pval \\ 
  \hline
(Intercept) & 3.85 & 0.09 & 42.21 & 0.00 \\ 
  test\_paramdensity & 0.01 & 0.12 & 0.07 & 0.94 \\ 
  test\_paramdotplot & 0.03 & 0.12 & 0.26 & 0.80 \\ 
  test\_paramhistogram & 0.23 & 0.12 & 1.97 & 0.05 \\ 
  d & -0.23 & 0.08 & -3.02 & 0.00 \\ 
  n1 & 0.00 & 0.00 & 1.74 & 0.08 \\ 
  I(n1/n2) & -0.04 & 0.02 & -1.48 & 0.14 \\ 
  test\_paramdensity:d & 0.11 & 0.11 & 1.02 & 0.31 \\ 
  test\_paramdotplot:d & 0.20 & 0.11 & 1.86 & 0.06 \\ 
  test\_paramhistogram:d & -0.01 & 0.11 & -0.09 & 0.93 \\ 
  test\_paramdensity:n1 & -0.00 & 0.00 & -2.15 & 0.03 \\ 
  test\_paramdotplot:n1 & -0.00 & 0.00 & -2.87 & 0.00 \\ 
  test\_paramhistogram:n1 & -0.00 & 0.00 & -0.93 & 0.35 \\ 
  test\_paramdensity:I(n1/n2) & 0.04 & 0.03 & 1.18 & 0.24 \\ 
  test\_paramdotplot:I(n1/n2) & 0.02 & 0.03 & 0.65 & 0.52 \\ 
  test\_paramhistogram:I(n1/n2) & 0.00 & 0.04 & 0.04 & 0.97 \\ 
   \hline
\end{tabular}
\end{center}
\caption{log time taken}

\end{table}

### 2012-03-28 setup.tex
small restructuring
p values in tables

Data was collected using Amazon's Mechanical Turk (MTurk) service. XXX More details: website and screen shot.
 Each participant was shown ten lineups. In order to avoid people `gaming the system' as found in earlier studies \cite{heer:2010, kosara:2010}, participants were shown two specially prepared lineups as a baseline of performance -- if participants got just one of these two reference lineups correct, we required at least an overall correct percentage of 20\%, which is indicative of a performance that is better than guessing. Table \ref{tbl:treatment} gives an overview of the number of times lineups in each combination of design and sample size were shown, and in how many of them the data plot was correctly identified.

@@

Besides correctness of responses, two additional response variables were recorded:  time taken for answer and a (subjective) confidence level, ranging from 1 (lowest) to 5 (highest), assessing how convinced participants are of the correctness of their answer. 
Figure \ref{fig:time} shows time taken until a decision was made in histograms, facetted by design and colored by correctness of answer. because of the skewness of the distributions, times were log-transformed. Polar charts take on average more time to answer, but are answered with much lower precision. Figure \ref{fig:conf} shows two barcharts of confidence levels by task, again coloring is used for correctness of answers. Euclidean displays lead to a very bimodal distribution of confidence: participants are either very sure or not sure at all of their answer. Confidence levels in polar charts are distributed much more uniformly.

@@

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=\linewidth]{conf-answer} 
   \caption{Barcharts of (subjective) confidence in answering correctly for each lineup. A strongly bimodal distribution is apparent, while confidence levels for polar charts are more uniform. There are only slight (and non significant) difference between confidence levels and correctness of results.}
   \label{fig:time}
\end{figure}

@@

\begin{table}[ht]
\begin{center}
\resizebox{\linewidth}{!} {
\rowcolors{2}{white}{lightgray}
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Error & $z$-value & $p$-value \\ 
  \hline
euclid & -0.08 & 0.39 & -0.21 & 0.84 \\ 
polar & -1.98 & 0.32 & -6.13 & 0.00 \\ [1pt]
  reflineTRUE & -0.14 & 0.26 & -0.53 & 0.59 \\ [1pt]
  sample\_size & 0.27 & 0.04 & 6.31 & 0.00 \\ [1pt]
  offset 90 & -0.43 & 0.37 & -1.18 & 0.24 \\ 
  offset 180 & -0.89 & 0.35 & -2.51 & 0.01 \\ 
  offset 270 & 0.21 & 0.38 & 0.55 & 0.58 \\ [1pt]
  polar:reflineTRUE & 0.51 & 0.35 & 1.44 & 0.15 \\ [1pt]
  polar:sample\_size & -0.23 & 0.05 & -5.02 & 0.00 \\ [1pt]
  polar:offset 90 & 0.64 & 0.49 & 1.30 & 0.20 \\ 
  polar:offset 180 & 0.91 & 0.47 & 1.92 & 0.05 \\ 
  polar:offset 270 & -0.73 & 0.54 & -1.35 & 0.18 \\
   \hline
\end{tabular}}
\end{center}
\caption{\label{tbl:correct} Model output for correct result. }
\end{table}

\begin{table}[ht]
\begin{center}
\resizebox{\linewidth}{!} {
\rowcolors{2}{white}{lightgray}
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & $z$ value & $p$ value from multcomp\\ 
  \hline
(Intercept) & 3.58 & 0.09 & 41.34 \\ 
  test\_parampolar & 0.22 & 0.10 & 2.17 \\ 
  reflineTRUE & 0.08 & 0.06 & 1.50 \\ 
  sample\_size & -0.03 & 0.00 & -8.73 \\ 
  offset90 & -0.06 & 0.08 & -0.72 \\ 
  offset180 & 0.16 & 0.08 & 2.07 \\ 
  offset270 & -0.04 & 0.07 & -0.54 \\ 
  test\_parampolar:reflineTRUE & -0.03 & 0.08 & -0.40 \\ 
  test\_parampolar:sample\_size & 0.03 & 0.01 & 6.25 \\ 
  test\_parampolar:offset90 & 0.14 & 0.11 & 1.23 \\ 
  test\_parampolar:offset180 & -0.17 & 0.11 & -1.55 \\ 
  test\_parampolar:offset270 & 0.17 & 0.11 & 1.50 \\ 
   \hline
\end{tabular}}
\end{center}
\caption{\label{tbl:time} Model output for (log) time taken. }
\end{table}

\begin{table}[ht]
\begin{center}
\resizebox{\linewidth}{!} {
\rowcolors{2}{white}{lightgray}
\begin{tabular}{rrrr}
  \hline
 & Estimate & Std. Error & t value \\ 
  \hline
(Intercept) & 3.58 & 0.09 & 41.34 \\ 
  test\_parampolar & 0.22 & 0.10 & 2.17 \\ 
  reflineTRUE & 0.08 & 0.06 & 1.50 \\ 
  sample\_size & -0.03 & 0.00 & -8.73 \\ 
  offset90 & -0.06 & 0.08 & -0.72 \\ 
  offset180 & 0.16 & 0.08 & 2.07 \\ 
  offset270 & -0.04 & 0.07 & -0.54 \\ 
  test\_parampolar:reflineTRUE & -0.03 & 0.08 & -0.40 \\ 
  test\_parampolar:sample\_size & 0.03 & 0.01 & 6.25 \\ 
  test\_parampolar:offset90 & 0.14 & 0.11 & 1.23 \\ 
  test\_parampolar:offset180 & -0.17 & 0.11 & -1.55 \\ 
  test\_parampolar:offset270 & 0.17 & 0.11 & 1.50 \\ 
   \hline
\end{tabular}
}
\end{center}
\caption{\label{tbl:confidence} Model output for confidence level. }
\end{table}

@@

Amazon's Mechanical Turk is an online service in which individuals can request and perform simple tasks and surveys for money. Using Turk we were able to gather data on how people responded to the charts. Each person is shown a series of ten charts and asked to choose the graphic on each chart which they think is "different". They then identify, from a list of choices, why they chose this particular chart as well as how confident they are on a scale of one to five. Personal information such as age group, gender, and education is also collected on a voluntary basis. The amount of time it took the individual to answer is also recorded. This study allows us to make informed judgments about how the plot type effects efficiency in terms of accuracy and time spent reading the plots. Xcite stephen few?X

>>>>

Data was collected using Amazon's Mechanical Turk (MTurk) service. XXX More details: website and screen shot.
 Each participant was shown ten lineups and asked to choose the graphic on each chart which they think is ``different". They then identify, from a list of choices, why they chose this particular chart as well as how confident they are on a scale of one to five. Personal information such as age group, gender, and education is collected on a voluntary basis. 
The amount of time it took the individual to answer is also recorded. This study allows us to make informed judgments about how design choices affect decoding \citet{cleveland:1994} in terms of accuracy and speed.

 In order to avoid people `gaming the system' as found in earlier studies \cite{heer:2010, kosara:2010}, participants were shown two specially prepared lineups as a baseline of performance -- if participants got just one of these two reference lineups correct, we required at least an overall correct percentage of 20\%, which is indicative of a performance that is better than guessing. Table \ref{tbl:treatment} gives an overview of the number of times lineups in each combination of design and sample size were shown, and in how many of them the data plot was correctly identified.

@@

%Besides correctness of responses, two additional response variables were recorded:  time taken for answer and a (subjective) confidence level, ranging from 1 (lowest) to 5 (highest), assessing how convinced participants are of the correctness of their answer. 
Figure \ref{fig:time} shows time taken until a decision was made in histograms, facetted by design and colored by correctness of answer. because of the skewness of the distributions, times were log-transformed. Polar charts take on average more time to answer, but are answered with much lower precision. Figure \ref{fig:conf} shows two barcharts of confidence levels by task, again coloring is used for correctness of answers. Euclidean displays lead to a very bimodal distribution of confidence: participants are either very sure or not sure at all of their answer. Confidence levels in polar charts are distributed much more uniformly.

@@

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=\linewidth]{conf-answer} 
   \caption{Barcharts of (subjective) confidence in answering correctly for each lineup. A strongly bimodal distribution is apparent, while confidence levels for polar charts are more uniform. There are only slight (and non significant) difference between confidence levels and correctness of results.}
   \label{fig:conf}
\end{figure}

@@

\begin{table}[ht]
\begin{center}
%\resizebox{\linewidth}{!} {
\rowcolors{2}{white}{lightgray}
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Error & $z$-value & $p$-value \\ 
  \hline
euclid & -0.08 & 0.39 & -0.21 & 0.84 \\ 
polar & -1.98 & 0.32 & -6.13 & 0.00 \\ [1pt]
  reflineTRUE & -0.14 & 0.26 & -0.53 & 0.59 \\ [1pt]
  sample\_size & 0.27 & 0.04 & 6.31 & 0.00 \\ [1pt]
  offset 90 & -0.43 & 0.37 & -1.18 & 0.24 \\ 
  offset 180 & -0.89 & 0.35 & -2.51 & 0.01 \\ 
  offset 270 & 0.21 & 0.38 & 0.55 & 0.58 \\ [1pt]
  polar:reflineTRUE & 0.51 & 0.35 & 1.44 & 0.15 \\ [1pt]
  polar:sample\_size & -0.23 & 0.05 & -5.02 & 0.00 \\ [1pt]
  polar:offset 90 & 0.64 & 0.49 & 1.30 & 0.20 \\ 
  polar:offset 180 & 0.91 & 0.47 & 1.92 & 0.05 \\ 
  polar:offset 270 & -0.73 & 0.54 & -1.35 & 0.18 \\
   \hline
\end{tabular}
%}
\end{center}
\caption{\label{tbl:correct} Model output for correct result. }
\end{table}

\begin{table}[ht]
\begin{center}
\resizebox{\linewidth}{!} {
\rowcolors{2}{white}{lightgray}
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std..Error & t.value & pval \\ 
  \hline
(Intercept) & 3.58 & 0.09 & 41.34 & 0.00 \\ 
  test\_parampolar & 0.22 & 0.10 & 2.17 & 0.03 \\ 
  reflineTRUE & 0.08 & 0.06 & 1.50 & 0.13 \\ 
  sample\_size & -0.03 & 0.00 & -8.73 & 0.00 \\ 
  offset90 & -0.06 & 0.08 & -0.72 & 0.47 \\ 
  offset180 & 0.16 & 0.08 & 2.07 & 0.04 \\ 
  offset270 & -0.04 & 0.07 & -0.54 & 0.59 \\ 
  test\_parampolar:reflineTRUE & -0.03 & 0.08 & -0.40 & 0.69 \\ 
  test\_parampolar:sample\_size & 0.03 & 0.01 & 6.25 & 0.00 \\ 
  test\_parampolar:offset90 & 0.14 & 0.11 & 1.23 & 0.22 \\ 
  test\_parampolar:offset180 & -0.17 & 0.11 & -1.55 & 0.12 \\ 
  test\_parampolar:offset270 & 0.17 & 0.11 & 1.50 & 0.13 \\ 
   \hline
\end{tabular}}
\end{center}
\caption{\label{tbl:time} Model output for (log) time taken. }
\end{table}

\begin{table}[ht]
\begin{center}
\resizebox{\linewidth}{!} {
\rowcolors{2}{white}{lightgray}
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std..Error & t.value & pval \\ 
  \hline
(Intercept) & 2.92 & 0.16 & 18.17 & 0.00 \\ 
  test\_parampolar & -0.30 & 0.16 & -1.86 & 0.06 \\ 
  reflineTRUE & -0.16 & 0.09 & -1.74 & 0.08 \\ 
  sample\_size & 0.01 & 0.01 & 1.01 & 0.31 \\ 
  offset90 & 0.14 & 0.13 & 1.07 & 0.28 \\ 
  offset180 & 0.03 & 0.13 & 0.26 & 0.80 \\ 
  offset270 & 0.01 & 0.12 & 0.07 & 0.94 \\ 
  test\_parampolar:reflineTRUE & 0.31 & 0.13 & 2.46 & 0.01 \\ 
  test\_parampolar:sample\_size & -0.01 & 0.01 & -0.70 & 0.49 \\ 
  test\_parampolar:offset90 & -0.14 & 0.18 & -0.78 & 0.44 \\ 
  test\_parampolar:offset180 & -0.01 & 0.18 & -0.04 & 0.97 \\ 
  test\_parampolar:offset270 & -0.00 & 0.18 & -0.01 & 0.99 \\ 
   \hline
\end{tabular}
}
\end{center}
\caption{\label{tbl:confidence} Model output for confidence level. }
\end{table}

@@

%Amazon's Mechanical Turk is an online service in which individuals can request and perform simple tasks and surveys for money. Using Turk we were able to gather data on how people responded to the charts. Each person is shown a series of ten charts and asked to choose the graphic on each chart which they think is "different". They then identify, from a list of choices, why they chose this particular chart as well as how confident they are on a scale of one to five. Personal information such as age group, gender, and education is also collected on a voluntary basis. The amount of time it took the individual to answer is also recorded. This study allows us to make informed judgments about how the plot type effects efficiency in terms of accuracy and time spent reading the plots. Xcite stephen few?X
